---
format:
  revealjs:
    theme: simple
    slide-number: true
    transition: fade
    center: false
execute:
  echo: false
  warning: false
  message: false
---
```{r}
#| label: setup
#| include: false

library(tidyverse)
library(knitr)
library(here)
library(kableExtra)

# Read raw data
retail_raw <- read_csv(
  here::here("data/Retail turnover, state by industry subgroup, seasonally adjusted.csv"),
  col_names = FALSE,
  show_col_types = FALSE
)

# Extract series names from first row
series_names <- retail_raw[1, -1] |>
  pivot_longer(cols = everything(), values_to = "series") |>
  pull(series)

# Clean and reshape data
retail_clean <- retail_raw[-(1:10), ] |>
  rename(Month = X1) |>
  pivot_longer(cols = -Month, names_to = "series_col", values_to = "Turnover") |>
  mutate(
    series_index = as.integer(str_remove(series_col, "X")) - 1,
    series = series_names[series_index],
    Month = as.Date(as.numeric(Month), origin = "1899-12-30"),
    Turnover = as.numeric(Turnover)
  ) |>
  separate(series, into = c("Measure", "State", "Industry"), sep = ";", fill = "right") |>
  mutate(
    across(c(Measure, State, Industry), str_trim),
    Industry = str_replace(
      Industry,
      "Furniture, floor coverings, houseware and textile goods retailing",
      "Furniture and houseware retailing"
    )
  ) |>
  filter(
    Month >= as.Date("2020-01-01"),
    Month <= as.Date("2025-05-01"),
    !str_detect(State, "Total"),
    !str_detect(Industry, "Total"),
    !is.na(Turnover)
  )
```
## Retail Sector Performance Across Australian States

### Identifying Growth Industries in the Post-Pandemic Era (2020–2025)

<br>

Shixu Xiao

Anali Nanayakkara Palliya Guruge

Tiansi Zhu

Sum Yee Chong

---

## Problem Introduction: Why Retail Recovery Matters

- COVID-19 substantially disrupted Australia's retail sector.

- Lockdowns and mobility restrictions changed consumer spending behaviour.

- Retail recovery has not occurred uniformly across Australian states and industries.

- Understanding these differences is important for businesses, investors, and policymakers.

---

## Research Focus

### Dataset

- Australian Bureau of Statistics (ABS) Retail Trade data

- Seasonally adjusted retail turnover

- January 2020 – May 2025

### Objective

Identify patterns of post-pandemic retail recovery across industries and regions.

## Methodology: Data Preparation {.smaller}

- Covers all **8 states and territories**, multiple industry subgroups
- CSV imported in wide format, pivoted to long format using `tidyverse`
- State and industry identifiers separated using `separate()`
- Aggregate totals and missing values excluded

```{r}
#| label: tbl-summary
#| tbl-cap: "Top 10 state-industry combinations by average monthly retail turnover, 2020–2025 (AUD millions, seasonally adjusted)"
#| echo: false

retail_clean |>
  group_by(State, Industry) |>
  summarise(average_turnover = round(mean(Turnover, na.rm = TRUE), 1), .groups = "drop") |>
  arrange(desc(average_turnover)) |>
  head(10) |>
  kable(col.names = c("State", "Industry", "Avg Turnover ($M)")) |>
  kableExtra::kable_styling(font_size = 11)
```
## Methodology: Analytical Approach {.smaller}

- Percentage change calculated from **January 2020 to May 2025** to measure growth
- Analysis focuses on both **state-level** and **industry-level** patterns
- Visualisations produced using `ggplot2`; tables rendered with `knitr`

```{r}
#| label: fig-trend
#| fig-cap: "Monthly retail turnover across selected industries in Australia, January 2020 to May 2025 (seasonally adjusted, AUD millions)"
#| echo: false
#| fig-width: 9
#| fig-height: 2.8

retail_clean |>
  filter(Industry %in% c(
    "Supermarket and grocery stores",
    "Cafes, restaurants and takeaway food services",
    "Clothing, footwear and personal accessory retailing"
  )) |>
  group_by(Month, Industry) |>
  summarise(Turnover = sum(Turnover), .groups = "drop") |>
  ggplot(aes(x = Month, y = Turnover, colour = Industry)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ Industry, scales = "free_y") +
  labs(x = "Date", y = "Turnover ($ Millions)") +
  theme_minimal() +
  theme(legend.position = "none", strip.text = element_text(size = 7))
```

## Results: Industry Growth Since 2020 {.smaller}

Food services led recovery; newspaper and book retailing continued structural decline.

```{r}
#| label: fig-growth
#| fig-cap: "Percentage change in total retail turnover by industry subgroup, January 2020 to May 2025 (seasonally adjusted, AUD millions). Industries ranked by growth rate."
#| echo: false
#| fig-width: 11
#| fig-height: 6.5

growth_data <- retail_clean |>
  filter(
    Month %in% as.Date(c("2020-01-01", "2025-05-01")),
    !Industry %in% c(
      "Food retailing",
      "Household goods retailing",
      "Clothing, footwear and personal accessory retailing",
      "Cafes, restaurants and takeaway food services",
      "Other retailing"
    )
  ) |>
  mutate(period = ifelse(Month == as.Date("2020-01-01"), "start", "end")) |>
  group_by(Industry, period) |>
  summarise(total = sum(Turnover, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = total) |>
  mutate(
    pct_change = ((end - start) / start) * 100,
    Industry = str_wrap(Industry, 35)
  ) |>
  filter(!is.na(pct_change))

ggplot(
  growth_data,
  aes(x = pct_change, y = reorder(Industry, pct_change), fill = pct_change > 0)
) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c")) +
  labs(x = "Percentage Change (%)", y = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.y = element_text(size = 10),
    panel.grid.major.y = element_blank()
  )
```
---
## Results: Key Takeaways {.smaller}

- Food services and health-related retail were the **primary drivers of post-pandemic growth**
- Newspaper and book retailing and department stores face **ongoing structural decline**
- Supermarket and grocery stores maintained **steady positive growth** throughout the period
- Hardware and building supplies showed **moderate growth**, driven by the home improvement boom
- Clear divide between **experience-based retail** (growing) and **traditional discretionary retail** (declining)
 
## Discussion & Conclusion {.smaller}
 
:::: {.columns}
 
::: {.column width="50%"}
 
**Discussion**
 
- **Food services** led growth at approx. 50% -- pent-up demand and shift to experience-based spending
- **Pharmaceutical and personal care retailing** grew steadily, driven by lasting health consciousness
- **Newspaper and book retailing** fell approx. 20% due to long-run digital substitution
- **Department stores** contracted under pressure from online retail
 
:::
 
::: {.column width="50%"}
 
**Conclusion**
 
- Recovery is **uneven but directionally clear**
- Food services and health-related retail are the **primary growth sectors**
- Print and department retail face **ongoing structural decline**
- **Victoria and WA** recorded the strongest state-level growth
 
:::
 
::::
 
## Recommendations {.smaller}
 
:::: {.columns}
 
::: {.column width="45%"}
 
```{r}
#| label: fig-state-growth-rec
#| fig-width: 4.5
#| fig-height: 3.8
#| echo: false
 
retail_clean |>
  filter(Month %in% as.Date(c("2020-01-01", "2025-05-01"))) |>
  mutate(period = ifelse(Month == as.Date("2020-01-01"), "start", "end")) |>
  group_by(State, period) |>
  summarise(total = sum(Turnover, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = total) |>
  mutate(pct_change = ((end - start) / start) * 100) |>
  ggplot(aes(x = pct_change, y = reorder(State, pct_change), fill = pct_change)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = "#abd9e9", high = "#2c7bb6") +
  labs(x = "Percentage Change (%)", y = NULL,
       title = "Growth by State (2020-2025)") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.y = element_text(size = 9),
    panel.grid.major.y = element_blank()
  )
```
 
:::
 
::: {.column width="55%"}
 
**For Businesses and Investors**
 
Prioritise food services, pharmaceutical and personal care sectors -- strongest and most sustained growth trends.
 
**For Policymakers**
 
Support declining sectors (newspaper/book retailing) with **transition assistance**, not recovery interventions -- decline is structural, not pandemic-driven.
 
**For Future Research**
 
Longitudinal analysis beyond 2025 to separate recovery effects from structural trends; state-level decomposition of growth drivers.
 
:::
 
::::
 
## References
 
Australian Bureau of Statistics (2025). *Retail Trade, Australia: Table 12 -- Retail Turnover, State by Industry Subgroup, Seasonally Adjusted.* Available at: https://www.abs.gov.au/statistics/industry/retail-and-wholesale-trade/retail-trade-australia/may-2025