---
format:
  revealjs:
    theme: simple
    slide-number: true
    transition: fade
    center: false
execute:
  echo: false
  warning: false
  message: false
---
```{r}
#| label: setup
#| include: false

library(tidyverse)
library(knitr)
library(here)
library(kableExtra)

# Read raw data
retail_raw <- read_csv(
  here::here("data/Retail turnover, state by industry subgroup, seasonally adjusted.csv"),
  col_names = FALSE,
  show_col_types = FALSE
)

# Extract series names from first row
series_names <- retail_raw[1, -1] |>
  pivot_longer(cols = everything(), values_to = "series") |>
  pull(series)

# Clean and reshape data
retail_clean <- retail_raw[-(1:10), ] |>
  rename(Month = X1) |>
  pivot_longer(cols = -Month, names_to = "series_col", values_to = "Turnover") |>
  mutate(
    series_index = as.integer(str_remove(series_col, "X")) - 1,
    series = series_names[series_index],
    Month = as.Date(as.numeric(Month), origin = "1899-12-30"),
    Turnover = as.numeric(Turnover)
  ) |>
  separate(series, into = c("Measure", "State", "Industry"), sep = ";", fill = "right") |>
  mutate(
    across(c(Measure, State, Industry), str_trim),
    Industry = str_replace(
      Industry,
      "Furniture, floor coverings, houseware and textile goods retailing",
      "Furniture and houseware retailing"
    )
  ) |>
  filter(
    Month >= as.Date("2020-01-01"),
    Month <= as.Date("2025-05-01"),
    !str_detect(State, "Total"),
    !str_detect(Industry, "Total"),
    !is.na(Turnover)
  )
```
## Retail Sector Performance Across Australian States

### Identifying Growth Industries in the Post-Pandemic Era (2020–2025)

<br>

Shixu Xiao

Anali Nanayakkara Palliya Guruge

Tiansi Zhu

Sum Yee Chong

---

## Problem Introduction: Why Retail Recovery Matters

- COVID-19 substantially disrupted Australia's retail sector.

- Lockdowns and mobility restrictions changed consumer spending behaviour.

- Retail recovery has not occurred uniformly across Australian states and industries.

- Understanding these differences is important for businesses, investors, and policymakers.

---

## Research Focus

### Dataset

- Australian Bureau of Statistics (ABS) Retail Trade data

- Seasonally adjusted retail turnover

- January 2020 – May 2025

### Objective

Identify patterns of post-pandemic retail recovery across industries and regions.

## Methodology: Data Preparation {.smaller}

- Covers all **8 states and territories**, multiple industry subgroups
- CSV imported in wide format, pivoted to long format using `tidyverse`
- State and industry identifiers separated using `separate()`
- Aggregate totals and missing values excluded

```{r}
#| label: tbl-summary
#| tbl-cap: "Top 10 state-industry combinations by average monthly retail turnover, 2020–2025 (AUD millions, seasonally adjusted)"
#| echo: false

retail_clean |>
  group_by(State, Industry) |>
  summarise(average_turnover = round(mean(Turnover, na.rm = TRUE), 1), .groups = "drop") |>
  arrange(desc(average_turnover)) |>
  head(10) |>
  kable(col.names = c("State", "Industry", "Avg Turnover ($M)")) |>
  kableExtra::kable_styling(font_size = 11)
```
## Methodology: Analytical Approach {.smaller}

- Percentage change calculated from **January 2020 to May 2025** to measure growth
- Analysis focuses on both **state-level** and **industry-level** patterns
- Visualisations produced using `ggplot2`; tables rendered with `knitr`

```{r}
#| label: fig-trend
#| fig-cap: "Monthly retail turnover across selected industries in Australia, January 2020 to May 2025 (seasonally adjusted, AUD millions)"
#| echo: false
#| fig-width: 9
#| fig-height: 2.8

retail_clean |>
  filter(Industry %in% c(
    "Supermarket and grocery stores",
    "Cafes, restaurants and takeaway food services",
    "Clothing, footwear and personal accessory retailing"
  )) |>
  group_by(Month, Industry) |>
  summarise(Turnover = sum(Turnover), .groups = "drop") |>
  ggplot(aes(x = Month, y = Turnover, colour = Industry)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ Industry, scales = "free_y") +
  labs(x = "Date", y = "Turnover ($ Millions)") +
  theme_minimal() +
  theme(legend.position = "none", strip.text = element_text(size = 7))
```



