---
title: "Retail Sector Performance Across Australian States: Identifying Growth Industries in the Post-Pandemic Era (2020–2025)"

author:
  - "Shixu Xiao"
  - "Anali Nanayakkara Palliya Guruge"
  - "Tiansi Zhu"
  - "Sum Chong"

format:
  typst:
    toc: true
    toc-depth: 2
    number-sections: true
    papersize: a4
    margin:
      x: 2.5cm
      y: 2.5cm

execute:
  echo: false
  warning: false
  message: false
  include: false

knitr:
  opts_chunk:
    fig.path: ../figures/

bibliography: references.bib
---

# Executive Summary

# Introduction
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
```

```{r}
retail_raw <- read_excel(
  "../data/Retail turnover, state by industry subgroup, seasonally adjusted.xlsx",
  sheet = "Data1",
  col_names = FALSE
)

retail_raw |> 
  slice(1:12)
```

```{r}
# Extract the series names from the first row of the Excel file
# Each column represents one retail time series
series_names <- retail_raw[1, -1] |>
  pivot_longer(
    cols = everything(),
    values_to = "series"
  ) |>
  pull(series)

# Clean and reshape the raw ABS retail dataset
retail_clean <- retail_raw[-(1:10), ] |>

  # Rename the first column as Month
  rename(Month = ...1) |>

  # Convert the data from wide format to long format
  pivot_longer(
    cols = -Month,
    names_to = "series_col",
    values_to = "Turnover"
  ) |>

  # Create useful variables and clean values
  mutate(

    # Extract the column index number
    series_index = as.integer(str_remove(series_col, "\\.\\.\\.")) - 1,

    # Match each column with its retail series description
    series = series_names[series_index],

    # Convert Excel numeric dates into R Date format
    Month = as.Date(as.numeric(Month), origin = "1899-12-30"),

    # Convert turnover values into numeric format
    Turnover = as.numeric(Turnover)
  ) |>

  # Split the series description into components
  separate(
    series,
    into = c("Measure", "State", "Industry"),
    sep = " ; ",
    fill = "right"
  ) |>

  # Remove unnecessary white spaces
  mutate(
    across(c(Measure, State, Industry), str_trim)
  ) |>

  # Restrict the analysis period to post-pandemic years
  filter(
    Month >= as.Date("2020-01-01"),
    Month <= as.Date("2025-05-01")
  )

# Inspect the cleaned dataset structure
glimpse(retail_clean)
```

```{r}
# Calculate total retail turnover by state
state_summary <- retail_clean |>
  group_by(State) |>
  summarise(
    Total_Turnover = sum(Turnover, na.rm = TRUE)
  ) |>
  arrange(desc(Total_Turnover))

# Display the state-level summary
state_summary
```

```{r}
# Calculate average turnover by state and year
state_growth <- retail_clean |>

  # Keep only the years used for comparison
  filter(year(Month) %in% c(2020, 2025)) |>

  # Create a year variable
  mutate(Year = year(Month)) |>

  # Calculate average turnover for each state and year
  group_by(State, Year) |>
  summarise(
    Avg_Turnover = mean(Turnover, na.rm = TRUE),
    .groups = "drop"
  ) |>

  # Convert the data from long to wide format
  pivot_wider(
    names_from = Year,
    values_from = Avg_Turnover,
    names_prefix = "Turnover_"
  ) |>

  # Calculate percentage growth between 2020 and 2025
  mutate(
    Growth_Percent =
      ((Turnover_2025 - Turnover_2020) /
         Turnover_2020) * 100
  ) |>

  # Rank states by growth rate
  arrange(desc(Growth_Percent))

# Display the state growth comparison
state_growth
```

```{r}
# Calculate average turnover by industry and year
industry_growth <- retail_clean |>

  # Keep only 2020 and 2025 observations
  filter(year(Month) %in% c(2020, 2025)) |>

  # Create a year variable
  mutate(Year = year(Month)) |>

  # Calculate average turnover for each industry and year
  group_by(Industry, Year) |>
  summarise(
    Avg_Turnover = mean(Turnover, na.rm = TRUE),
    .groups = "drop"
  ) |>

  # Convert the data into wide format
  pivot_wider(
    names_from = Year,
    values_from = Avg_Turnover,
    names_prefix = "Turnover_"
  ) |>

  # Calculate percentage growth
  mutate(
    Growth_Percent =
      ((Turnover_2025 - Turnover_2020) /
         Turnover_2020) * 100
  ) |>

  # Rank industries by growth rate
  arrange(desc(Growth_Percent))

# Display industry growth comparison
industry_growth
```

The Australian retail sector experienced substantial disruption during the COVID-19 pandemic, followed by uneven recovery patterns across states and retail industries. Changes in consumer behaviour, reopening policies, labour market conditions, and inflation contributed to significant variation in retail turnover between 2020 and 2025. Understanding which industries recovered most strongly is important for businesses, policymakers, and investors seeking to identify emerging growth opportunities in the post-pandemic economy.

This report analyses seasonally adjusted retail turnover data published by the Australian Bureau of Statistics (ABS), covering multiple Australian states and retail industry subgroups. Using reproducible workflows implemented in Quarto and R, the report examines differences in retail performance across regions and industries during the post-pandemic recovery period.

Preliminary analysis suggests that retail recovery patterns differed substantially across states, with Victoria and Western Australia recording some of the strongest growth rates between 2020 and 2025. Industry-level analysis indicates particularly strong growth in cafes, restaurants, takeaway food services, clothing retailing, and personal accessory retailing. These findings provide insight into structural changes in Australian consumer spending and highlight sectors that demonstrated strong post-pandemic resilience.

# Methodology

# Results

# Discussion and Conclusion

# References