
title: "Retail Sector Performance Across Australian States: Identifying Growth Industries in the Post-Pandemic Era (2020–2025)"

author:
  - "Shixu Xiao"
  - "Anali Nanayakkara Palliya Guruge"
  - "Tiansi Zhu"
  - "Sum Yee Chong"


format:
  typst:
    toc: true
    toc-depth: 2
    number-sections: true
    papersize: a4
    margin:
      x: 2.5cm
      y: 2.5cm

execute:
  echo: false
  warning: false
  message: false
  include: false

knitr:
  opts_chunk:
    fig.path: ../figures/

bibliography: references.bib
---

# Executive Summary

# Introduction


The Australian retail sector experienced substantial disruption during the COVID-19 pandemic, followed by uneven recovery patterns across states and retail industries. Changes in consumer behaviour, reopening policies, labour market conditions, and inflation contributed to significant variation in retail turnover between 2020 and 2025. Understanding which industries recovered most strongly is important for businesses, policymakers, and investors seeking to identify emerging growth opportunities in the post-pandemic economy.

This report analyses seasonally adjusted retail turnover data published by the Australian Bureau of Statistics (ABS), covering multiple Australian states and retail industry subgroups. Using reproducible workflows implemented in Quarto and R, the report examines differences in retail performance across regions and industries during the post-pandemic recovery period.

Preliminary analysis suggests that retail recovery patterns differed substantially across states, with Victoria and Western Australia recording some of the strongest growth rates between 2020 and 2025. Industry-level analysis indicates particularly strong growth in cafes, restaurants, takeaway food services, clothing retailing, and personal accessory retailing. These findings provide insight into structural changes in Australian consumer spending and highlight sectors that demonstrated strong post-pandemic resilience.

# Methodology
Monthly retail turnover data were obtained from the Australian Bureau of Statistics (ABS) Retail Trade, Australia release (Cat. No. 8501.0, Table 12), which tracks seasonally adjusted turnover figures broken down by state and industry subgroup. This publication is updated monthly and freely accessible via abs.gov.au, with the current release extending through May 2025, making it a reliable source for examining recent retail trends across Australia.

The dataset captures turnover in millions of Australian dollars across all eight states and territories — New South Wales, Victoria, Queensland, South Australia, Western Australia, Tasmania, the Northern Territory, and the Australian Capital Territory — and spans thirteen retail industry subgroups, ranging from supermarkets and grocery stores to cafes, clothing retailers, and household goods outlets. Given the study's focus on post-pandemic recovery, only observations from January 2020 onward were retained, providing a window that encompasses both the acute disruption phase of COVID-19 and the subsequent recovery period through to May 2025.

The raw data arrived in wide format, with each column encoding a distinct state-industry combination. Column labels were extracted from the first row and reassigned before the table was pivoted into long format via the tidyverse package in R; state and industry identifiers were then isolated from the original series strings using the separate() function. National aggregate rows and records with missing values were excluded to ensure the analysis reflected state-level patterns only. Industry labels were also simplified where necessary to improve readability.

All visualisations were produced using ggplot2 and tables were rendered with knitr. @tbl-summary lists the ten state-industry combinations with the highest average monthly turnover, offering a snapshot of which segments consistently dominated retail activity. @fig-trend traces how turnover evolved across three key industry subgroups between 2020 and 2025, highlighting divergent recovery trajectories in the post-pandemic period.
```{r}
#| label: data-cleaning

library(readxl)
library(tidyverse)
library(knitr)

retail_names <- read_excel(
  "../data/Retail turnover, state by industry subgroup, seasonally adjusted.xlsx",
  sheet = "Data1",
  n_max = 1,
  col_names = FALSE
)
```
```{r}
retail_raw <- read_excel(
  "../data/Retail turnover, state by industry subgroup, seasonally adjusted.xlsx",
  sheet = "Data1",
  skip = 10,
  col_names = FALSE
)
```
```{r}
names(retail_raw) <- as.character(retail_names[1, ])
names(retail_raw)[1] <- "Date"
retail_raw <- retail_raw[, !is.na(names(retail_raw))]
```

```{r}
retail_long <- retail_raw |>
  pivot_longer(
    cols = -Date,
    names_to = "series",
    values_to = "turnover"
  )
```

```{r}
retail_clean <- retail_long |>
  separate(
    series,
    into = c("measure", "state", "industry"),
    sep = ";"
  ) |>
  mutate(
    Date = as.Date(Date),
    state = str_trim(state),
    industry = str_trim(industry),
    industry = str_replace(
      industry,
      "Furniture, floor coverings, houseware and textile goods retailing",
      "Furniture and houseware retailing"
    )
  ) |>
  filter(
    Date >= as.Date("2020-01-01"),
    Date <= as.Date("2025-05-01"),
    state != "Total (State)",
    industry != "Total (Industry)",
    !is.na(turnover)
  )
```
```{r}
#| label: tbl-summary
#| tbl-cap: "Top 10 state-industry combinations by average monthly retail turnover, 2020–2025 (AUD millions, seasonally adjusted)"


retail_clean |>
  group_by(state, industry) |>
  summarise(
    average_turnover = round(mean(turnover, na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  arrange(desc(average_turnover)) |>
  head(10) |>
  kable(
    col.names = c("State", "Industry", "Average Turnover ($M)")
  )
```

```{r}
#| label: fig-trend
#| fig-cap: "Monthly retail turnover across selected industries in Australia, January 2020 to May 2025 (seasonally adjusted, AUD millions)"
#| fig-width: 10
#| fig-height: 4

retail_clean |>
  filter(
    industry %in% c(
      "Supermarket and grocery stores",
      "Cafes, restaurants and takeaway food services",
      "Clothing, footwear and personal accessory retailing"
    )
  ) |>
  group_by(Date, industry) |>
  summarise(
    turnover = sum(turnover),
    .groups = "drop"
  ) |>
  ggplot(aes(x = Date, y = turnover, colour = industry)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ industry, scales = "free_y") +
  labs(
    x = "Date",
    y = "Turnover ($ Millions)",
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8)
  )
```


# Results

# Discussion and Conclusion


# References

