---
title: "Retail Sector Performance Across Australian States: Identifying Growth Industries in the Post-Pandemic Era (2020–2025)"

author:
  - "Shixu Xiao"
  - "Anali Nanayakkara Palliya Guruge"
  - "Tiansi Zhu"
  - "Sum Yee Chong"

format:
  typst:
    toc: true
    toc-depth: 2
    number-sections: true
    papersize: a4
    margin:
      x: 2.5cm
      y: 2.5cm

execute:
  echo: false
  warning: false
  message: false
  include: false

knitr:
  opts_chunk:
    fig.path: figures/

bibliography: references.bib
---

# Executive Summary

This report examines post-pandemic retail sector performance across Australian states and industries using seasonally adjusted Australian Bureau of Statistics (ABS) retail turnover data from January 2020 to May 2025. The analysis shows that food services, pharmaceutical retailing, and personal care sectors recorded the strongest growth, while newspaper and book retailing and department stores experienced continued decline. State-level results indicate that Victoria and Western Australia were among the strongest-performing regions during the recovery period. These findings suggest that businesses and investors should prioritise high-growth service and health-related retail sectors, while policymakers should focus on supporting structural adjustment in declining industries.

# Introduction

The Australian retail sector experienced substantial disruption during the COVID-19 pandemic as lockdowns, mobility restrictions, and changing consumer behaviour affected spending patterns across the country. Although public health restrictions have largely been removed, retail recovery has not occurred uniformly across states and industry sectors. Understanding these differences is important for businesses seeking growth opportunities, investors evaluating retail performance, and policymakers supporting economic recovery.

Retail turnover is a widely used indicator of consumer spending activity and provides valuable insights into the health of the retail sector. This report investigates retail sector performance across Australian states and retail industry subgroups during the post-pandemic period. The analysis uses seasonally adjusted retail turnover data published by the Australian Bureau of Statistics (ABS), providing a reliable measure of retail activity over time. The dataset covers all Australian states and territories and includes a broad range of retail industries, such as food retailing, household goods retailing, clothing retailing, and food services.

A reproducible analytical workflow was implemented using Quarto and R to ensure transparency, consistency, and replicability throughout the project. The primary objective of the report is to examine retail sector performance across Australian states and industries between January 2020 and May 2025 and to identify patterns of post-pandemic recovery. The remainder of the report presents the data preparation process, summarises the main findings, and discusses their implications for businesses, investors, and policymakers.

# Methodology

Monthly retail turnover data were obtained from the Australian Bureau of Statistics (ABS) Retail Trade, Australia release (Cat. No. 8501.0, Table 12), which tracks seasonally adjusted turnover figures broken down by state and industry subgroup. This publication is updated monthly and freely accessible via abs.gov.au, with the current release extending through May 2025, making it a reliable source for examining recent retail trends across Australia.

The dataset captures turnover in millions of Australian dollars across all eight states and territories — New South Wales, Victoria, Queensland, South Australia, Western Australia, Tasmania, the Northern Territory, and the Australian Capital Territory — and spans multiple retail industry subgroups, ranging from supermarkets and grocery stores to cafes, clothing retailers, and household goods outlets. Only observations from January 2020 onward were retained, providing a window that encompasses both the acute disruption phase of COVID-19 and the subsequent recovery period through to May 2025.

The raw data were imported as a CSV file in wide format, with each column encoding a distinct state-industry combination. Column labels were extracted from the first row using index-based matching before the table was pivoted into long format via the tidyverse package in R; state and industry identifiers were then isolated using the separate() function. Aggregate industry totals and records with missing values were excluded to ensure the analysis reflected state-level patterns only. Industry labels were also simplified where necessary to improve readability.

All visualisations were produced using ggplot2 and tables were rendered with knitr. @tbl-summary lists the ten state-industry combinations with the highest average monthly turnover, offering a snapshot of which segments consistently dominated retail activity. @fig-trend traces how turnover evolved across supermarket and grocery stores, cafes and restaurants, and clothing retailing between 2020 and 2025, highlighting divergent recovery trajectories in the post-pandemic period.

```{r}
#| label: data-cleaning

library(readxl)
library(tidyverse)
library(knitr)
library(here)

# Read raw data without column names
retail_raw <- read_csv(
  here::here("data/Retail turnover, state by industry subgroup, seasonally adjusted.csv"),
  col_names = FALSE,
  show_col_types = FALSE
)

# Extract series names from first row
series_names <- retail_raw[1, -1] |>
  pivot_longer(
    cols = everything(),
    values_to = "series"
  ) |>
  pull(series)

# Clean and reshape data
retail_clean <- retail_raw[-(1:10), ] |>
  rename(Month = X1) |>
  pivot_longer(
    cols = -Month,
    names_to = "series_col",
    values_to = "Turnover"
  ) |>
  mutate(
    series_index = as.integer(str_remove(series_col, "X")) - 1,
    series = series_names[series_index],
    Month = as.Date(as.numeric(Month), origin = "1899-12-30"),
    Turnover = as.numeric(Turnover)
  ) |>
  separate(
    series,
    into = c("Measure", "State", "Industry"),
    sep = ";",
    fill = "right"
  ) |>
  mutate(
    across(c(Measure, State, Industry), str_trim),
    Industry = str_replace(
      Industry,
      "Furniture, floor coverings, houseware and textile goods retailing",
      "Furniture and houseware retailing"
    )
  ) |>
  filter(
    Month >= as.Date("2020-01-01"),
    Month <= as.Date("2025-05-01"),
    !str_detect(State, "Total"),
    !str_detect(Industry, "Total"),
    !is.na(Turnover)
  )
```

```{r}
#| label: tbl-summary
#| tbl-cap: "Top 10 state-industry combinations by average monthly retail turnover, 2020–2025 (AUD millions, seasonally adjusted)"
#| include: true

retail_clean |>
  group_by(State, Industry) |>
  summarise(
    average_turnover = round(mean(Turnover, na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  arrange(desc(average_turnover)) |>
  head(10) |>
  kable(
    col.names = c("State", "Industry", "Average Turnover ($M)")
  )
```

```{r}
#| label: fig-trend
#| fig-cap: "Monthly retail turnover across selected industries in Australia, January 2020 to May 2025 (seasonally adjusted, AUD millions)"
#| fig-width: 10
#| fig-height: 4
#| include: true

retail_clean |>
  filter(
    Industry %in% c(
      "Supermarket and grocery stores",
      "Cafes, restaurants and takeaway food services",
      "Clothing, footwear and personal accessory retailing"
    )
  ) |>
  group_by(Month, Industry) |>
  summarise(
    Turnover = sum(Turnover),
    .groups = "drop"
  ) |>
  ggplot(aes(x = Month, y = Turnover, colour = Industry)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ Industry, scales = "free_y") +
  labs(
    x = "Date",
    y = "Turnover ($ Millions)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8)
  )
```

# Results

@fig-growth presents the percentage change in total retail turnover by industry subgroup between January 2020 and May 2025, summed across all Australian states.

Cafes, restaurants and catering services recorded the strongest growth over the period, consistent with the recovery of dining and hospitality following COVID-19 restrictions. Takeaway food services also showed strong gains, reflecting sustained demand for convenient dining options. Pharmaceutical, cosmetic and toiletry goods retailing grew steadily, driven by continued consumer spending on health and personal care.

In contrast, newspaper and book retailing recorded the sharpest decline, continuing a long-run structural trend driven by digital substitution. Department stores also contracted, reflecting ongoing pressure from online retail and shifting consumer preferences.

Hardware, building and garden supplies showed moderate growth, likely supported by the pandemic-era home improvement boom. Supermarket and grocery stores maintained steady positive growth throughout the period, underpinned by consistent household demand.

Overall, food services and health-related retail were the primary drivers of post-pandemic growth, while print-based and discretionary retail continued to decline.

```{r}
#| label: fig-growth
#| fig-cap: "Percentage change in total retail turnover by industry subgroup, January 2020 to May 2025 (seasonally adjusted, AUD millions). Industries ranked by growth rate."
#| fig-width: 8
#| fig-height: 5
#| include: true

growth_data <- retail_clean |>
  filter(
    Month %in% as.Date(c("2020-01-01", "2025-05-01")),
    !Industry %in% c(
      "Food retailing",
      "Household goods retailing",
      "Clothing, footwear and personal accessory retailing",
      "Cafes, restaurants and takeaway food services",
      "Other retailing"
    )
  ) |>
  mutate(period = ifelse(Month == as.Date("2020-01-01"), "start", "end")) |>
  group_by(Industry, period) |>
  summarise(total = sum(Turnover, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = total) |>
  mutate(
    pct_change = ((end - start) / start) * 100,
    Industry = str_wrap(Industry, 35)
  ) |>
  filter(!is.na(pct_change))

ggplot(
  growth_data,
  aes(x = pct_change, y = reorder(Industry, pct_change), fill = pct_change > 0)
) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("TRUE" = "#2c7bb6", "FALSE" = "#d7191c")) +
  labs(x = "Percentage Change (%)", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    panel.grid.major.y = element_blank()
  )
```

# Discussion, conclusion and recommendations

**Discussion**

The results reveal a clear bifurcation in Australian retail performance between 2020 and 2025. Food services — particularly cafes, restaurants, and takeaway — emerged as the strongest growth segment, as shown in Figure 2, with cafes and restaurants recording approximately 50% growth over the period. This recovery reflects pent-up consumer demand following prolonged COVID-19 dining restrictions, as well as a structural shift toward experience-based spending once lockdowns lifted. The sustained growth in pharmaceutical, cosmetic, and toiletry goods retailing similarly reflects a lasting behavioural legacy of the pandemic, with heightened consumer attention to personal health and hygiene driving ongoing demand.

In contrast, newspaper and book retailing recorded the sharpest decline of approximately 20%, continuing a long-run structural trend driven by digital substitution that predates the pandemic. Department stores also contracted, reflecting compounding pressures from the growth of online retail platforms and shifting consumer preferences away from traditional discretionary retail formats. These segments are unlikely to recover without significant business model transformation.

The dominance of New South Wales, Victoria, and Queensland in absolute turnover (Table 1) is consistent with their population size. However, Western Australia's relative strength in growth rates warrants attention, potentially reflecting its earlier economic recovery and resource-sector-driven income growth during this period.

**Conclusion**

As shown in @fig-state-growth, Victoria and Western Australia recorded the strongest growth rates among all states.

This analysis of seasonally adjusted ABS retail turnover data from January 2020 to May 2025 demonstrates that post-pandemic retail recovery in Australia has been uneven but directionally clear. Food services and health-related retail are the primary growth industries, while print media retail and traditional department stores face ongoing structural decline. Victoria and Western Australia recorded some of the strongest state-level growth, suggesting regional variation in recovery pace driven by differing reopening timelines and economic conditions.

```{r}
#| label: fig-state-growth
#| fig-cap: "Percentage change in total retail turnover by state, January 2020 to May 2025 (seasonally adjusted, AUD millions)."
#| fig-width: 7
#| fig-height: 4
#| include: true

retail_clean |>
  filter(Month %in% as.Date(c("2020-01-01", "2025-05-01"))) |>
  mutate(period = ifelse(Month == as.Date("2020-01-01"), "start", "end")) |>
  group_by(State, period) |>
  summarise(total = sum(Turnover, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = total) |>
  mutate(pct_change = ((end - start) / start) * 100) |>
  ggplot(aes(x = pct_change, y = reorder(State, pct_change), fill = pct_change)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = "#abd9e9", high = "#2c7bb6") +
  labs(x = "Percentage Change (%)", y = NULL) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    panel.grid.major.y = element_blank()
  )
```

**Recommendations**

For Businesses and Investors: Capital allocation should prioritise food services, pharmaceutical retailing, and personal care sectors when considering retail expansion in Australia, as these segments demonstrate the strongest and most sustained post-pandemic growth trajectories.

For Policymakers: Support programs targeting structurally declining sectors such as newspaper and book retailing should account for the long-run digital substitution trend rather than treating decline as a purely pandemic-driven phenomenon. Targeted transition assistance may be more effective than recovery-focused interventions.

For Future Research: Longitudinal analysis extending beyond 2025 would help distinguish pandemic recovery effects from underlying structural trends. State-level decomposition of growth drivers — separating population growth, real income effects, and behavioural change — would provide more granular policy-relevant insights.

# References

Data were obtained from the Australian Bureau of Statistics [@abs_retail_2025].
